\documentclass[10pt]{report}
\usepackage{/Users/bradenhoagland/latex/math}

\lhead{Braden Hoagland}
\chead{HW 9}
\rhead{}

\renewcommand{\theenumi}{\alph{enumi}}

\begin{document}
%\tableofcontents

{\color{blue}Problems completed: All.}

\begin{exer}[\S 51 \#1]
If $h,h':X\to Y$ are homotopic and $k,k':Y\to Z$ are homotopic, then $k \circ h$ and $k' \circ h'$ are homotopic.
\end{exer}
{\color{blue}Collaborators: None.}

Suppose $h \simeq h'$ via  $H$ and $k \simeq k'$ via $K$, then we claim that $k \circ h \simeq k'\circ h'$ via the map
\begin{align*}
	F: X \times I &\to Z\\
	(x,t)&\mapsto K(H(x,t),t).
\end{align*}
At $t=0$, $F(x,0) = K(H(x,0),0) = K(h(x),0) = (k\circ h)(x)$. Similarly, at $t=1$, $F(x,1)=K(H(x,1),1)=K(h'(x),1) = (k'\circ h')(x)$. Finally, $F$ is continuous since it is the composition of continuous functions: Let $H' = (H, \pi_2)$, then $H'$ is continuous and $F = K \circ H'$.

\pagebreak
\begin{exer}[\S 51 \#3]
	$X$ is \textbf{contractible} if $i_{X}$ is nulhomotopic.
	\begin{enumerate}
		\item $I=[0,1]$ and $\mathbb{R}$ are contractible.
		\item A contractible space is path connected.
		\item If $Y$ is contractible, then for any $X$, the set $[X,Y]$ (the set of homotopy classes of maps of $X$ into $Y$) has a single element.
		\item If $X$ is contractible and $Y$ is path connected, then $[X,Y]$ has a single element.
	\end{enumerate}
\end{exer}
{\color{blue}Collaborators: None.}

\begin{enumerate}
	\item If a space is convex, then any two paths are homotopic via the straight line homotopy. Since $I$ and $\mathbb{R}$ are both convex, they are both contractible.

	\item Suppose $X$ is contractible and $x,y \in X$. Since $X$ is contractible, $i_{X}$ and some constant function $\tilde{f}(x) = c$ are homotopic via some $F$. Since $F$ is continuous by definition,
		\begin{align*}
			f_{x} &\cdot F(x,\cdot) \\
			f_{y} &\cdot F(y,\cdot)
		\end{align*}
		are both continuous. Then $f_{x}$ is a path from $x$ to $c$, and $f_{y}$ is a path from $y$ to $c$. Then $f:[0,2]\to X$ given by
		\[
			f(t) \doteq
			\begin{cases}
				f_{x}(t) & t \leq 1 \\
				f_{y}(2-t) & t \geq 1
			\end{cases}
		\] is a path from $x$ to $y$. Thus $X$ is path connected.

	\item Since $Y$ is contractible, $i_{Y}$ and some constant function $y\mapsto c$ are homotopic via some $G$. Let $f:X\to Y$ be arbitrary, then $F(x,t) = G(f(x),t)$ is a homotopy from  $f$ to $\tilde{f}$, where $\tilde{f}(x)=c$. Since $f$ was arbitrary, this means that all maps from $X$ to $Y$ are homotopic to $\tilde{f}$. Since homotopy is an equivalence relation, this means all such maps are homotpic to each other, i.e. $[X,Y]$ has one element.

	\item Fix $y \in Y$. Since $X$ is contractible, $i_{X}$ and some constant function $x \mapsto c$ are homotopic via some $G$. Then for all maps $f:X\to Y$, the map $F(x,t) = f(G(x,t))$ is a homotopy from $f$ to the constant function $x \mapsto f(c)$. Now since $Y$ is path connected, there is a path $\gamma$ from $f(c)$ to $y$. Then $H(x,t) = \gamma(t)$ is a homotopy from $x \mapsto f(c)$ to $x \mapsto y$. Composing these two homotopies shows $f \simeq x \mapsto y$. But since $f$ was arbitrary, this means all maps from $X$ to $Y$ are homotopic to the same constant function. Since homotopy is an equivalence relation, this means $[X,Y]$ has one element.
\end{enumerate}

\end{document}
