\documentclass[10pt]{report}
\usepackage{/Users/bradenhoagland/latex/math}

\lhead{Braden Hoagland}
\chead{Math 502 First Day Survey}
\rhead{}

\begin{document}
%\tableofcontents

\textbf{Name:} Braden Hoagland.

\textbf{501:} I didn't take 501, but we did communicate over the break about this. I took 401 with Professor Jiuya Wang.

\textbf{Favorite math classes:} Probably 531. I'm taking my first pure math electives this semester (411 and 421), though, so I'm hoping I can add a few more classes to that list soon.

\textbf{Best part about winter break:} I slept in every day and finally had time to study things for fun instead of just for class. I was also around my family every day, which was a nice change of pace from living with just a roommate.

\textbf{Thursday Availability:} Yes, I'm available at 3:15 on Thursdays.

\textbf{Confidence Ratings:}
\begin{itemize}
	\item Group Theory: 4. I reviewed this pretty in-depth over the break.
	\item Ring Theory: 3. I reviewed this over the break, but not as in-depth.
	\item Module Theory: 1. I didn't know what module theory was before I signed up for this course.
	\item Latex: 5. I use Latex a lot.
\end{itemize}

\textbf{Algebraic topics:} The topic I'm most interested in now would probably be Galois theory, but we're already covering that in this class. I've been wanting to study universal algebra, but am unsure what I would even need to know first before I could start.

\textbf{Other comments:} I'm pretty nervous heading into the class because I didn't have the prereq course, so I wouldn't be surprised if I spend a lot of time in office hours trying to improve my understanding of the material.

\end{document}
